The Influence of Radial Drift and Gas Accretion on the C/O Ratio in Protoplanetary Disks

### 2016-04-13 Dissertate-Harvard-LaTeX/chapters/chapter3.tex
upd

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentclass[12pt, preprint]{aastex}
\documentclass[apj]{emulateapj}

\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex

%\usepackage{pdfsync}
\usepackage{subeqnarray}
\usepackage{natbib}
\usepackage{color}
\usepackage[utf8]{inputenc}

\bibliographystyle{apj}

\newcommand{\ie}{i.e.\ }
\newcommand{\eg}{e.g.\ }
\newcommand{\p}{\partial}
\newcommand{\brak}[1]{\langle #1\rangle}

\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
\newcommand{\cm}{\; {\rm cm}}
\newcommand{\mm}{\; {\rm mm}}
%\newcommand{\ps}{\; {\rm s^{-1}}}
\newcommand{\km}{\; {\rm km}}
\newcommand{\au}{\; \varpi_{\rm AU}}
\newcommand{\AU}{\; {\rm AU}}
\def\K{\; {\rm K}}

\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
\newcommand{\nab}{\vc{\nabla}}
\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}

\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
\newcommand{\App}[1]{Appendix~\ref{#1}}
\newcommand{\Sec}[1]{Sect.~\ref{#1}}
\newcommand{\Chap}[1]{Chapter~\ref{#1}}
\newcommand{\Fig}[1]{Fig.~\ref{#1}}
\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
\newcommand{\Tab}[1]{Table \ref{#1}}

\definecolor{gray}{gray}{0.5}
\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}

%\newenvironment{packed_item}{
%\begin{itemize}
%  \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%}{\end{itemize}}

\begin{document}

%\slugcomment{Draft Modified \today}

\title{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}

\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Tilman Birnstiel\altaffilmark{1, 3}, Ruth A. Murray-Clay\altaffilmark{2}}
\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\altaffiltext{2}{Department of Physics, University of California, Santa Barbara, CA 93106}
\altaffiltext{3}{Max Planck Institute for Astronomy, K\"onigstuhl 17, 69117 Heidelberg, Germany}

\begin{abstract}
%The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry. 
The C/O ratio is a defining feature of both gas giant atmospheric and protoplanetary disk chemistry. In disks, the C/O ratio is regulated by the presence of snowlines of major volatiles at different distances from the central star. We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk. We determine the snowline locations for a range of fixed initial particle sizes and disk types. For our fiducial disk model, we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb at a size-dependent location in the disk, which is independent of the particle's initial position. The snowline radius decreases for larger particles, up to sizes of $\sim$7~m. Compared to a static disk, we find that radial drift and gas accretion in a viscous disk move the H$_2$O snowline inwards by up to 40 \%, the CO$_2$ snowline by up to 60 \%, and the CO snowline by up to 50 \%. We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for. 
\end{abstract}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{drift_timescales_betaS1_gas_acc_new2.pdf}
%\vspace{-0.5in}
\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines). The timescales are calculated at three representative locations, i.e. the H$_2$O, CO$_2$ and CO snowlines in the static disk. For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively. The horizontal dot-dashed line represents a typical disk lifetime of 3 Myr. The particle size ordering at the minimum $t_{\rm drift}$ is not monotonic in snowline distance due to different drag regimes for those particle sizes at the snowline locations (Epstein drag at the H$_2$O and CO$_2$ snowlines, and Stokes drag at the CO snowline). Similarly, the ordering of $t_{\rm des}$ is not monotonic in snowline distance due to the non-monotony in mean molecular weight between H$_2$O, CO$_2$ and CO (18 $m_{\rm p}$, 44 $m_{\rm p}$ and 28 $m_{\rm p}$, respectively). Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e. $t_{\rm drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.} 
\label{fig:timescales}
\end{figure}

@@

\begin{figure*}[tb]
\centering
\includegraphics[width=0.7\textwidth]{desorption_distance_passive_active_colorbar_test2.pdf}
%\vspace{-0.5in}
\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels). The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines). The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar. For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.} 
\label{fig:snowlines}
\end{figure*}

\begin{figure*}[t!]
\centering
\includegraphics[width=0.7\textwidth]{desorption_distance_actual_vs_estimated_passive_active_new.pdf}
%\vspace{-0.5in}
\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text). The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).  The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
\label{fig:an_vs_actual}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{s_t_a.pdf}
%\vspace{-0.5in}
\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in an irradiated disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left). Particles desorb almost instantaneously. Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel. Particles lose most of their mass very close to the distance at which they fully desorb. The static H$_2$O snowline is shown for reference (dashed vertical lines).}
\label{fig:s_t_a}
\end{figure*}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.55\textwidth]{C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
%\vspace{-0.5in}
\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel).
%, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
The particle size increases from $\sim$0.05 cm to $\sim$700 cm as indicated by the color bar. The horizontal dotted line represents the stellar value of 0.54. The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}) for the bottom panel. 
%The snowline location moves inward as the particle size increases. 
%The curves exterior to the static snowline represent desorption distances in the absence of readsorption.  %Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline. 
For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, as shown by the arrows.}
%, with the decrease more significant interior to the H$_2$O snowline.}
\label{fig:CO_ratio}
\end{figure}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{time_plots.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve). For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
\label{fig:timeplots}
\end{figure*}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{desorption_distance_varying_Md.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines). The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve). A lower disk mass does not change the snowline location.}
\label{fig:varMd}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{desorption_distance_transition_disk_1000.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines). The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}. Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
\label{fig:cavity}
\end{figure}

@@

\acknowledgements{We thank the referee for a helpful and detailed report. We also thank Sean Andrews for useful conversations and suggestions. T. B. is grateful for support from the NASA Origins of Solar Systems grant NNX12AJ04G and the Smithsonian Institution Pell Grant program.}

\appendix
\section{Steady-state active disk solution} \label{app:steadystate}

Following \citet{shakura73} and \citet{armitage10}, the steady-state solution for a geometrically thin, optically thick actively accreting disk with an $\alpha$-prescription for viscosity is governed by the following set of equations:

\begin{subeqnarray}
\label{eq:diskeq}
\nu & = & \alpha c H \slabel{eq:nueq} \\
c^2 & = & \frac{k_{\rm B} T_{\rm act}}{\mu m_{\rm p}} \slabel{eq:cdeq} \\
\rho & = & \frac{1}{\sqrt{2 \pi}} \frac{\Sigma}{H} \slabel{eq:rhoeq} \\
H & = & \frac{c}{\Omega_{k}} \slabel{eq:Heq} \\
T_{\rm act}^4 & = & \frac{3}{4} \tau T_{\rm surf}^4 \slabel{eq:Teq} \\
\tau & = & \frac{1}{2} \Sigma \kappa \slabel{eq:taueq} \\
\nu \Sigma & = & \frac{\dot{M}}{3 \pi} \slabel{eq:Mdot} \\
\sigma T_{\rm surf}^4 & = & \frac{9}{8} \nu \Sigma \Omega_{\rm k}^2 \slabel{eq:nusigeq} \\
\kappa & = & \kappa_0 T_{\rm act}^2 \slabel{eq:kappaeq},
\end{subeqnarray}
where $T_{\rm surf}$ is the surface temperature of the disk and the other quantities are defined in the main text. This is a system of nine equations with nine unknowns ($\nu$, $c$, $H$, $T_{\rm act}$, $\rho$, $\Sigma$, $\tau$, $T_{\rm surf}$, $\kappa$) that can be solved numerically once $\alpha$ and $\kappa_0$ are specified.

\section{Desorption distance analytic solution} \label{app:tdriftan}

For a particle of size $s$ that desorbs and satisfies $\tau_{\rm s} \ll 1$ ($\tau_{\rm s}$ is the dimensionless stopping time, defined in Section \ref{sec:drift}), we can derive an explicit analytic solution for the particle's desorption distance in an irradiated disk. For $\tau_{\rm s} \ll 1$, a particle is in the Epstein drag regime (see Equation \ref{eq:ts}) and its drift velocity $\dot{r}$ (Equation \ref{eq:rdotpas}) can be approximated as
\begin{equation}
\label{eq:rdotapp}
\dot{r} \approx -2 \eta \Omega_{\rm k} r \tau_{\rm s}.
\end{equation}
By using Equations (\ref{eq:tdes}) and (\ref{eq:tdrift}) and setting $t_{\rm drift}=t_{\rm des}$, we can express a particle's desorption distance as

\begin{equation}
\label{eq:tan}
r_{\rm des}=\Bigg(\frac{d}{q \,C} \,\, \mathcal{W}\Big[\frac{(B/A)^{-q/d} q\,C}{d}\Big]\Bigg)^{\frac{1}{q}},
\end{equation}
where $\mathcal{W}$ is the Lambert-W function, $q=3/7$ is the power-law coefficient in Equation (\ref{eq:diskT}), $d=-\frac{1}{2}+p-q$ with $p=1$ the power-law coefficient in Equation (\ref{eq:disksigma}), and 
\begin{subeqnarray}
%\label{eq:tan}
A & = &  \frac{\rho_0}{\rho_{\rm s}} \frac{r_0^2}{s c_0} r_0^d \\ 
B & = & \frac{\rho_{\rm s} s}{3 \mu_{\rm x} N_{\rm x} \nu_{\rm x}} \\
C & = & \frac{E_{\rm x}}{T_0} r_0^{-q}, \\
\end{subeqnarray}
where $r_0=1$ AU, $\rho_0=\rho(r_0)$ and $c_0=c(r_0)$.

%
%\emgr{Right now it's unclear to me what could go in an appendix, if anything. Maybe discuss a bit the algorithm to evolve $\Sigma_{\rm p}$ (although it is already explained in detail in the appendix of Birnstiel et al. 2010). Maybe show some example profiles of $\Sigma_{\rm p}$ at different times and for different particle sizes.}

\if\bibinc n
\bibliography{refs}
\fi

\if\bibinc y
\begin{thebibliography}
\end{thebibliography}
\fi

\end{document}

>>>>

%\begin{savequote}[75mm]
%Nulla facilisi. In vel sem. Morbi id urna in diam dignissim feugiat. Proin molestie tortor eu velit. Aliquam erat volutpat. Nullam ultrices, diam tempus vulputate egestas, eros pede varius leo.
%\qauthor{Quoteauthor Lastname}
%\end{savequote}

\chapter{C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion}

\section*{abstract}
%The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry. 
The C/O ratio is a defining feature of both gas giant atmospheric and protoplanetary disk chemistry. In disks, the C/O ratio is regulated by the presence of snowlines of major volatiles at different distances from the central star. We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk. We determine the snowline locations for a range of fixed initial particle sizes and disk types. For our fiducial disk model, we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk, and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving and viscous disk, desorb at a size-dependent location in the disk, which is independent of the particle's initial position. The snowline radius decreases for larger particles, up to sizes of $\sim$7~m. Compared to a static disk, we find that radial drift and gas accretion in a viscous disk move the H$_2$O snowline inwards by up to 40 \%, the CO$_2$ snowline by up to 60 \%, and the CO snowline by up to 50 \%. We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{figures/drift_timescales_betaS1_gas_acc_new2.pdf}
%\vspace{-0.5in}
\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines). The timescales are calculated at three representative locations, i.e. the H$_2$O, CO$_2$ and CO snowlines in the static disk. For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively. The horizontal dot-dashed line represents a typical disk lifetime of 3 Myr. The particle size ordering at the minimum $t_{\rm drift}$ is not monotonic in snowline distance due to different drag regimes for those particle sizes at the snowline locations (Epstein drag at the H$_2$O and CO$_2$ snowlines, and Stokes drag at the CO snowline). Similarly, the ordering of $t_{\rm des}$ is not monotonic in snowline distance due to the non-monotony in mean molecular weight between H$_2$O, CO$_2$ and CO (18 $m_{\rm p}$, 44 $m_{\rm p}$ and 28 $m_{\rm p}$, respectively). Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e. $t_{\rm drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.} 
\label{fig:timescales}
\end{figure}

@@

\begin{figure*}[tb]
\centering
\includegraphics[width=0.7\textwidth]{figures/desorption_distance_passive_active_colorbar_test2.pdf}
%\vspace{-0.5in}
\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels). The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines). The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar. For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.} 
\label{fig:snowlines}
\end{figure*}

\begin{figure*}[t!]
\centering
\includegraphics[width=0.7\textwidth]{figures/desorption_distance_actual_vs_estimated_passive_active_new.pdf}
%\vspace{-0.5in}
\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text). The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).  The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
\label{fig:an_vs_actual}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{figures/s_t_a.pdf}
%\vspace{-0.5in}
\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in an irradiated disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left). Particles desorb almost instantaneously. Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel. Particles lose most of their mass very close to the distance at which they fully desorb. The static H$_2$O snowline is shown for reference (dashed vertical lines).}
\label{fig:s_t_a}
\end{figure*}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.55\textwidth]{figures/C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
%\vspace{-0.5in}
\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel).
%, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
The particle size increases from $\sim$0.05 cm to $\sim$700 cm as indicated by the color bar. The horizontal dotted line represents the stellar value of 0.54. The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}) for the bottom panel. 
%The snowline location moves inward as the particle size increases. 
%The curves exterior to the static snowline represent desorption distances in the absence of readsorption.  %Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline. 
For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, as shown by the arrows.}
%, with the decrease more significant interior to the H$_2$O snowline.}
\label{fig:CO_ratio}
\end{figure}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{figures/time_plots.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve). For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
\label{fig:timeplots}
\end{figure*}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{figures/desorption_distance_varying_Md.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines). The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve). A lower disk mass does not change the snowline location.}
\label{fig:varMd}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{figures/desorption_distance_transition_disk_1000.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines). The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}. Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
\label{fig:cavity}
\end{figure}

@@
